A Bayesian approach to burial networks to infer social changes in 17th northeastern Taiwan during the European colonization period

### 2021-01-02 analysis/paper/paper.Rmd
added one reference and citation
change title and revise the way for making ties

---
title: "Bayesian exponential random graph modeling of KWL Burial network"
author:
  - Author One
  - Author Two
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  keyword 1; keyword 2; keyword 3
highlights: |
  These are the highlights. 
---

@@

<!-- burial data and the definition of tie formation  --> 
We analyzed burial data collected from the published excavation reports, and the original fieldwork notes of Kiwulan. Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of three levels, high, medium, and low, according to their distributions across all burials. Gold-foil beads are in levels of high (>3), medium (=2-3), and low (=1); glass beads and carnelian beads are in levels of high (>6), medium (=3-6), and low (=1-2). That means if burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them. For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods in common (i.e. presence or absence). Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and burial value index. Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting according to historical records. The burial value index is an important attribute for economical inequality that we assigned the burials into three classes, high (>30), medium (>11 and <30), and low (<12), as an index of wealth. The number is the sum of values of each type of prestige good, which is calculate by the total number of a prestige item from burial context over the number of prestige item in a burial. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.

@@

<!-- comparison between two networks using bootstrap --> 
One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials. To understand the robustness of comparison between two networks, we used vertex bootstrap to cross-validate the results based on Bayesian network analysis. Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e. node) of a network to quantify standard errors in the network statistics of interest [@Chen2019; @Snijders1999]. This enables the evaluation of uncertainty for networks and tests the difference between multiple networks by examining their confidence intervals for the network population. Networks with unoverlapped intervals means there is a significant difference. We set the resampling size to 1,000 networks to compute endogenous network statistics, including density, transitivity, and popularity for our two networks. The result shows a significant difference in network popularity between the two networks, which is consistent with our finding of negative popularity in the pre-European period and positive popularity in the post-European period using the Bayesian approach. For density and transitivity, the 95% confidence intervals from two networks overlapped with each other that indicates no significant difference. This is also consistent with our results of Bayesian modeling, where both networks present similar positive or negative effects some degree of difference.

>>>>

---
title: "A Bayesian approach to burial networks to infer social changes in 17th northeastern Taiwan during the European colonization period"
author:
  - Author One
  - Author Two
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: "../templates/template.docx" # Insert path for the DOCX file
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  Bayesian modeling; network analysis; Monte-carlo simulation; Burials; European colonialism; social differentiation; Taiwan
highlights: |
  These are the highlights. 
---

@@

<!-- burial data and the definition of tie formation  --> 
We analyzed burial data collected from the published excavation reports, and the original fieldwork notes of Kiwulan. Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials. Gold-foil beads are in levels of high (>3), upper-middle (=3), lower-middle (=2), and low (=1); carnelian beads are in levels of high (>6), upper-middle (=4-6), lower-middle (=3), and low (=1-2); glass beads are in levels of high (>6), upper-middle (=3-6), lower-middle (=2), and low (=1). That means if burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them. For less frequent prestige goods, including imported porcelains, gold foils, and fish-shaped ornaments, we linked two burials when they possess each type of goods in common (i.e. presence or absence). Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and burial value index. Ritual pottery was identified as locally made ceramics placed above graves that suggests funeral feasting according to historical records. The burial value index is an important attribute for economical inequality that we assigned the burials into three classes, high (>30), medium (>11 and <30), and low (<12), as an index of wealth. The number is the sum of values of each type of prestige good, which is calculate by the total number of a prestige item from burial context over the number of prestige item in a burial. Since burials tend to have multiple prestige goods in common, the network ties are weighted instead of binary data (where the value 1 represents a tie and the value 0 otherwise) [@Snijders2011]. For example, if two burials have both low quantity of glass beads and porcelain in common, the tie is given a value of 2. Our networks are non-directed, which means ties have no orientation forming the relationship between actors to indicate a mutual relationship.

@@

<!-- comparison between two networks using bootstrap --> 
One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials. To understand the robustness of comparison between two networks, we used vertex bootstrap to cross-validate the results based on Bayesian network analysis. Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e. node) that is useful for quantifying standard errors and estimating sampling variability in the network statistics of interest [@Chen2019; @Snijders1999; @Rroberts2021]. This enables the evaluation of uncertainty for networks and tests the difference between multiple networks by examining their confidence intervals for the network population. Networks with unoverlapped intervals means there is a significant difference. We set the resampling size to 1,000 networks to compute endogenous network statistics, including density, transitivity, and popularity for our two networks. The result shows a significant difference in network popularity between the two networks, which is consistent with our finding of negative popularity in the pre-European period and positive popularity in the post-European period using the Bayesian approach. For density and transitivity, the 95% confidence intervals from two networks overlapped with each other that indicates no significant difference. This is also consistent with our results of Bayesian modeling, where both networks present similar positive or negative effects some degree of difference.
